### 2011-04-03 msm2aa.tex
Started msm2aa, made title.tex, modified format.tex

\documentclass[12pt]{article}
\input{./include/format.tex}

% SET THIS STUFF - info for front page.
\newcommand{\nameOfTheModule}{MSM2template}
\newcommand{\nameOfTheAuthor}{Will Ridgers}
\newcommand{\nameOfTheLecturer}{Dr Qianxi Wang}

\begin{document}
\input{./include/title.tex}
\tableofcontents
\pagebreak

\section{Functions of Several Real Variables}
	\begin{defn} 
		For a function $z = f(x,y)$ a vertical section if the graph of $z = f(x,c)$ or $z = f(c,y)$ 
		for some constant $c$. A level curve is the curve $c = f(x,y)$.
	\end{defn}

\section{Partial derivatives and chain rule}

	\begin{defn}
		The function $f: \mathbb{R}^3 \to \mathbb{R}$ is said to have a partial derivative with respect to $x$
		at the point $(x_0, y_0, z_0)$ if there exists a limit
			\[
				\frac{\partial f (x_0, y_0, z_0)}{\partial x} \equiv \lim_{\Delta x \to 0} 
				\frac{f(x_0 + \Delta x, y_0, z_0) - f(x_0, y_0, z_0)}{\Delta x}
			\]
		which is called the partial derivative of $f$ with respect to $x$.
	\end{defn}

	Derivatives with respect to $y$ and $z$ are defined in a similar way.

	\begin{defn}
		The function $f(x,y,z)$ is called differentiable at $(x_0, y_0, z_0)$ if 
		$\Delta f = f(x,y,z) - f(x_0, y_0, z_0)$ can be expressed as
		\[
			\Delta f = f_x(x_0, y_0, z_0) \Delta x + f_y(x_0, y_0, z_0) \Delta y + f_z(x_0, y_0, z_0) \Delta z + o(\rho)
		\]
		where $\Delta x = x - x_0$, $\Delta y = y - y_0$, $\Delta z = z - z_0$ and 
		\[
			\rho = \sqrt{ (\Delta x)^2 + (\Delta y)^2 + (\Delta z)^2 }
		\]
	\end{defn}
	The notation $o(\rho)$ means going to $0$ faster than $\rho$. Note, this is infinitely small so can be ignored.

\section{Vector and vector geometry}
\section{Gradient vector and directional derivative}

\end{document}