### 2011-04-04 msm2aa.tex
50% of msms2aa done. no gfx.

\documentclass[12pt]{article}
\input{./include/include.tex}

% SET THIS STUFF - info for front page.
\newcommand{\nameOfTheModule}{MSM2Aa}
\newcommand{\nameOfTheAuthor}{Will Ridgers}
\newcommand{\nameOfTheLecturer}{Dr Qianxi Wang}

\begin{document}
\input{./include/title.tex}
\tableofcontents
\pagebreak

\section{Functions of Several Real Variables}
	\subsection{Functions of Several Real Variables}
		\begin{defn} 
			For a function $z = f(x,y)$ a vertical section if the graph of $z = f(x,c)$ or $z = f(c,y)$ 
			for some constant $c$. A level curve is the curve $c = f(x,y)$.
		\end{defn}

	\subsection{Partial derivatives and chain rule}

		\subsubsection{Partial derivatives}
		\begin{defn}
			The function $f: \mathbb{R}^3 \to \mathbb{R}$ is said to have a partial derivative with respect to $x$
			at the point $(x_0, y_0, z_0)$ if there exists a limit
				\[
					\frac{\partial f (x_0, y_0, z_0)}{\partial x} \equiv \lim_{\Delta x \to 0} 
					\frac{f(x_0 + \Delta x, y_0, z_0) - f(x_0, y_0, z_0)}{\Delta x}
				\]
			which is called the partial derivative of $f$ with respect to $x$.
		\end{defn}

		Derivatives with respect to $y$ and $z$ are defined in a similar way.

		\subsubsection{Differentiable functions}
		\begin{defn}
			The function $f(x,y,z)$ is called differentiable at $(x_0, y_0, z_0)$ if 
			$\Delta f = f(x,y,z) - f(x_0, y_0, z_0)$ can be expressed as
			\[
				\Delta f = f_x(x_0, y_0, z_0) \Delta x + f_y(x_0, y_0, z_0) \Delta y + f_z(x_0, y_0, z_0) \Delta z + o(\rho)
			\]
			where $\Delta x = x - x_0$, $\Delta y = y - y_0$, $\Delta z = z - z_0$ and 
			\[
				\rho = \sqrt{ (\Delta x)^2 + (\Delta y)^2 + (\Delta z)^2 }
			\]
		\end{defn}
		The notation $o(\rho)$ means going to $0$ faster than $\rho$. Note, this is infinitely small so can be ignored.

		\subsubsection{Sufficiently smooth}
		\begin{defn}
			(Informal) A function is said to be sufficiently smooth if the function and as many of it's partial derivatives
			as required are continuous where they need to be.
		\end{defn}

		\subsubsection{Chain rule}
		\begin{defn}
			Chain Rule for two variables. Let $f$ be a function of the variables $x_1, x_2, \cdots , x_n$, $f = f(x_1, x_2, \cdots , x_n)$ 
			where each $x_j$ is a function of the variables $t_1, t_2, \cdots , t_n$, $x_j = x_j(t_1, t_2, \cdots , t_n)$. Provided
			that $f$ and $x_j$ are sufficiently smooth, then
			\[
				\frac{\partial f}{\partial t_i} = \frac{\partial f}{\partial x_1}\frac{\partial x_1}{\partial t_i} + 
												\frac{\partial f}{\partial x_2}\frac{\partial x_2}{\partial t_i} + 
												\cdots +
												\frac{\partial f}{\partial x_n}\frac{\partial x_n}{\partial t_i}
			\]
		\end{defn}

	\subsection{Vector and vector geometry}

		\subsubsection{Vectors}
		\begin{defn}
			A vector has both a magnitude and a direction. Vectors may be denoted as $\mathbf{u}$, $\vec{u}$, or $\underline{u}$.
			The magnitude of a vector is denoted by 
			\[
				|\mathbf{u}| = \sqrt{u_1^2 + u_2^2 + u_3^2}
			\]
			A unit vector has length $1$ and in the Cartesian coordinate system $\mathbf{i}$, $\mathbf{j}$, and $\mathbf{k}$ denote the the unit 
			coordinate vectors along x, y, z-axes respectively.
		\end{defn}

		\subsubsection{Scalar (dot) product}
		\begin{defn}
			Given two vectors $\mathbf{a} = a_1 \mathbf{i} + a_2 \mathbf{j} + a_3 \mathbf{k}$,
			$\mathbf{b} = b_1 \mathbf{i} + b_2 \mathbf{j} + b_3 \mathbf{k}$ the scalar (or dot) product of the two is
			\[
				\mathbf{a} \cdot \mathbf{b} = |\mathbf{a}||\mathbf{b}|\cos{\theta} = a_1 b_1 + a_2 b_2 + a_3 b_3
			\]
			where $\theta$ is the angle between $\mathbf{a}$ and $\mathbf{b}$.
		\end{defn}

		\subsubsection{Vector (cross) product}
		\begin{defn}
			Given two vectors $\mathbf{a} = a_1 \mathbf{i} + a_2 \mathbf{j} + a_3 \mathbf{k}$,
			$\mathbf{b} = b_1 \mathbf{i} + b_2 \mathbf{j} + b_3 \mathbf{k}$ the vector (or cross) product of the two is
			\[
				\mathbf{a} \times \mathbf{b} = |\mathbf{a}||\mathbf{b}|\sin{\theta}\mathbf{n} =
				\begin{vmatrix}
					\mathbf{i} & \mathbf{j} & \mathbf{k} \\
					a_1 & a_2 & a_3 \\
					b_1 & b_2 & b_3 
				\end{vmatrix} 
			\]
		\end{defn}

		\subsubsection{Straight line equation}
		\begin{defn}
			The vector equation of a straight line is
			\[
				\mathbf{r} = \mathbf{a} + t \mathbf{d}
			\]
			for $t \in \mathbb{R}$, where $\mathbf{a}$ is a point on the line, $\mathbf{d}$ is parallel to the line, 
			and $\mathbf{r}$ is the position vector of a general point on the line.
		\end{defn}

		\subsubsection{3 dimensional plane equation}
		\begin{defn}
			The vector equation of a plane in 3 dimensions is
			\[
				\mathbf{r} \cdot \mathbf{n} = \mathbf{a} \cdot \mathbf{n}
			\]
			where $\mathbf{a}$ is the position vector of a point in the plane, $\mathbf{n}$ is a vector normal to
			the plane and $\mathbf{r}$ is the position vector of a general point in the plane.
		\end{defn}

		\subsubsection{Parametric form}
		A curve $C$ can be expressed using parametric form
		\[
			\mathbf{r}(t) = x(t)\mathbf{i} + y(t)\mathbf{j} + z(t)\mathbf{k}
		\]
		where $a \leq t \leq b$. Thus the initial point $A$ is $(x(a),y(a),z(a))$ and the terminal point $B$ at $(x(b),y(b),z(b))$.
		$\mathbf{r}(t)$ is a mapping on the interval $[a,b]$ on the t-axis to the curbe in xyz-space.

		\subsubsection{Tangent vector of a curve}
		For a curve represented in parametric form (see above), for each $t$ there is a point $\mathbf{r}(t)$ on the curve. The tangent
		vector of the curve can be found by
		\begin{align*}
			\frac{d\mathbf{r}}{dt}& = \lim_{\Delta t \to 0} \frac{\mathbf{r}(t+\Delta t) - \mathbf{r}(t)}{\Delta t} \\
							& = \lim_{\Delta t \to 0} \frac{x(t+\Delta t) - x(t)}{\Delta t} \mathbf{i} + 
								  \lim_{\Delta t \to 0} \frac{y(t+\Delta t) - y(t)}{\Delta t} \mathbf{j} + 
								  \lim_{\Delta t \to 0} \frac{z(t+\Delta t) - z(t)}{\Delta t} \mathbf{k} \\
							& = \frac{dx}{dt} \mathbf{i} + \frac{dy}{dt} \mathbf{j} + \frac{dz}{dt} \mathbf{k}
		\end{align*} 
		The unit tangent vector $\tau$ is $\tau \equiv \frac{\mathbf{r}_t}{|\mathbf{r}_t|}$. In general $\mathbf{r}' \not{=} 0$ for
		a sufficiently smooth function.

	\subsection{Gradient vector and directional derivative}

		\subsubsection{Gradient vector (grad/nabla/del)}
		\begin{defn}
			Let $f:\mathbb{R}^3 \to \mathbb{R}$ be a function. If the partial derfivatives of $f$ with respect to $x$, $y$, and $z$
			at the point $(a,b,c)$ then the gradient vector of $f$ at $(a,b,c)$ is defined to be the vector
			\[
				\nabla f(a,b,c) = \frac{\partial f(a,b,c)}{\partial x}\mathbf{i} + 
									\frac{\partial f(a,b,c)}{\partial y}\mathbf{j} + 
									\frac{\partial f(a,b,c)}{\partial z}\mathbf{k}
			\]
		\end{defn}

		\subsubsection{Key points}
		\begin{itemize}
			\item The value of $\nabla f$ at a point is independent of the coordinate system.
			\item $\nabla f$ is a vector function in n dimensions where f has n variables.
			\item $\nabla$ is a linear operator, i.e $\nabla (\lambda f + \mu g) = \lambda \nabla f + \mu \nabla g$.
		\end{itemize}

		\subsubsection{Grad as normal}
		\begin{thm}
			Let $f:\mathbb{R}^2 \to \mathbb{R}$ be a sufficiently function, and $(a,b)$ be a point on the level curve $f(x,y)=k$ for some
			constant $k$. If $\nabla f (a,b) \not{=} 0$, then the vector $\nabla f (a,b)$ is normal to the level curbe $f(x,y)=k$ at the point $(a,b)$.
		\end{thm}

		\subsubsection{Finding tangent plane to a surface}
		To find the tangent plane to a surface $z=F(x,y)$ or $G(x,y,z)=k$ at the point $\mathbf{a} = (a,b,c)$:
		\begin{enumerate}
			\item Write the surface in the form of the level surface
				\[
					f(x,y,z) = F(x,y) - z = 0
				\]
				or
				\[
					f(x,y,z) = G(x,y,z) - k = 0
				\]
			\item Calculate the normal vector to the surface at the point $\mathbf{a}$
				\[
					\mathbf{n} = \nabla f ( \mathbf{a} )
				\]
			\item The tangent plane is
				\[
					\mathbf{r} \cdot \mathbf{n} = \mathbf{a} \cdot \mathbf{n}
				\]
		\end{enumerate}

		\subsubsection{Directional derivatives}
		\begin{defn}
			Suppose that $f:\mathbb{R}^3 \to \mathbb{R}$ and $\mathbf{u} = u_1 \mathbf{i} + u_2 \mathbf{j} + u_3 \mathbf{k}$ is a unit vector in $\mathbb{R}^3$.
			The directional derivative of $f$ at $(x,y,z)$ in the direction of $\mathbf{u}$ is defined to be
			\[
				D_u f(x,y,z) = \lim_{h \to 0^{+}} \frac{f(x+hu_1, y+hu_2, z+hu_3) - f(x,y,z)}{h}
			\]
		\end{defn}

		\begin{lemma}
			Let $f:\mathbb{R}^3 \to \mathbb{R}$ be a sufficiently smooth, and $\mathbf{u} = u_1 \mathbf{i} + u_2 \mathbf{j} + u_3 \mathbf{k}$ 
			be a unit vector in $\mathbb{R}^3$. If it exists, then
			\[
				\frac{d}{dt} f(a + tu_1, b + tu_2, c + tu_3) | _ t=0 = D_u f(a,b,c)
			\]
		\end{lemma}

		\subsubsection{Relation between gradient and directional derivative}
		\begin{thm}
			Let $f:\mathbb{R}^3 \to \mathbb{R}$ be a sufficiently smooth, and $\mathbf{u} = u_1 \mathbf{i} + u_2 \mathbf{j} + u_3 \mathbf{k}$ 
			be a unit vector in $\mathbb{R}^3$. Then
			\[
				D_u f(a,b,c) = \mathbf{u} \cdot \nabla f(a,b,c)
			\]
		\end{thm}

		\begin{itemize}
			\item The directional derivative is a scalar function not a vector.
			\item $D_u$ is linear; $D_u(\lambda f + \mu g) = \lambda D_u f + \mu D_u g$.
		\end{itemize}

		\subsubsection{Calculating directional derivatives}
		To calculate the directional derivatibe of a function $f$ in the direction of the vector $\mathbf{u}$:
		\begin{enumerate}
			\item Find the unit vector $\hat{\mathbf{u}} = \frac{\mathbf{u}}{|\mathbf{u}|}$
			\item Find $\nabla f$
			\item Calculate $D_{\hat{\mathbf{u}}} f = \nabla f \cdot \hat{\mathbf{u}}$
		\end{enumerate}

		\subsubsection{Steepst ascent theorem}
		\begin{thm}
			Let $f$ be a function of several real variables. Suppose that $\nabla f$ exists and is non-zero at the point $\mathbf{a}$. 
			Then the direction of the vector $\nabla f(\mathbf{a})$ is always the direction of maximum increase of the function $f$ at the point $\mathbf{a}$.
		\end{thm}

		\subsubsection{Finding paths of maximum increase}
		To calculate the path of maximum increase of a function $f$ of two variables starting at point $(a,b)$:
		\begin{enumerate}
			\item Let the path be $\mathbf{r}(t) = x(t)\mathbf{i} + y(t)\mathbf{j}$
			\item $\nabla f$ at a point $(x,y)$ is always in the direction of the maximum increase (by Steepst ascent theorem) of the function $f$ at the point,
					so the tangent vector to the path  $\mathbf{r}'(t) = x'(t)\mathbf{i} + y'(t)\mathbf{j}$ is parallel to $\nabla f = f_x \mathbf{i} + f_y \mathbf{j}$
			\item Since the two vectors are parallel we have
					\[
						x'(t) \mathbf{i} + y'(t) \mathbf{j} = \lambda ( f_x \mathbf{i} + f_y \mathbf{j} ),  \quad \lambda \not{=} 0 \implies \frac{y'(t)}{x'(t)} = \frac{f_y}{f_x}
					\]
			\item Hence
					\[
						\frac{dy}{dx} = \frac{y'(t)}{x'(t)} = \frac{f_y}{f_x}
					\]
			\item Solve for $y$ in terms of $x$ using the boundary conditions $y=b$ when $x=a$.
		\end{enumerate}

\section{Integration of Functions of Several Real Variables}

	\subsection{Multiple integrals}

\end{document}