### 2011-04-05 msm2ba.tex
Started 2ca (Will) and 2ba (Joao)

\documentclass[12pt]{article}
\input{./include/include.tex}

% SET THIS STUFF - info for front page.
\newcommand{\nameOfTheModule}{MSM2Ba}
\newcommand{\nameOfTheAuthor}{Jo\~{a}o Cho\c{c}a}
\newcommand{\nameOfTheLecturer}{Dr Neal Bez}

\begin{document}
\input{./include/title.tex}
\tableofcontents
\pagebreak

\section{Assumed Knowledge}
	\subsection{Intervals}
		\begin{defn}
			An interval is a subset of $\mathbb{R}$ taking one of the following forms, where $a$ and $b$ are arbitrary real numbers and $a < b$.

			\begin{itemize}
				\item $(a,b) := \{x \in \mathbb{R}: a < x < b\}$ (this is called an open interval)
				\item $(a,b] := \{x \in \mathbb{R}: a < x \leq b\}$	
				\item $[a,b) := \{x \in \mathbb{R}: a \leq x < b\}$		
				\item $[a,b] := \{x \in \mathbb{R}: a \leq x \leq b\}$ (this is called a closed interval)
				\item $(a,\infty) := \{x \in \mathbb{R}: x > a\}$		
				\item $[a,\infty) := \{x \in \mathbb{R}: x \geq a\}$		
				\item $(-\infty,b) := \{x \in \mathbb{R}: x < b\}$		
				\item $(-\infty,b] := \{x \in \mathbb{R}: x \leq b\}$
				\item $(-\infty, \infty) := \mathbb{R}$
				\item The interval $(0, \infty)$ will also be denoted by $\mathbb{R}^+$.
			\end{itemize}

		\end{defn}

\end{document}